Improvement of Korean Morphological Analysis System \\Through Transformer-Based Re-Ranking

### 2023-12-02 contents/draft.tex
Apply revision by soomgo: Introduction

    Korean morphological analysis is the process of determining parts of speech by identifying morphemes, which are the smallest units of linguistic expression with independent meanings in a sentence.
    In an isolating language, such as English, this identification can be achieved relatively easily by tagging parts of speech sequentially.
    However, in Korean, the nature of the agglutinative language requires separating endings or postpositions and restoring inflections to their original form.
    In addition, because the basic input of other Korean analysis tasks is often a separate morpheme, the accuracy of the morphological analysis significantly affects the performance of Korean analysis.
    Modern high-performance deep learning methods in natural language processing (NLP) use a tokenization process that breaks the text into smaller units and converts each token into a vector as an input to the computational model~\cite{Mikolov2013}.
    Here, the token unit is mainly a subword unit, and to reflect the characteristics of a Korean subword, tokenization with separate morphemes is attempted in advance~\cite{SongHJ2021}.
    Using the results of the morphological analysis for this tokenization process improves the overall performance of the analysis by reflecting the semantic units of Korean.
    This requires a highly accurate and fast morphological analyzer.

    Various approaches have been proposed for morphological analysis, which is crucial in Korean language analysis~\cite{KwonHC1991, LeeDG2009, ShimKS2011, LeeJS2011, ShinJC2012, LeeCK2013, NaSH2014, NaSH2015, HwangHS2016, KimHM2016, ChungES2016, LeeCH2016, Li2017, NaSH2018, KimSW2018, ChoiYS2018, MinJW2018, MinJW2019, KimHM2019, SongHJ2019, MinJW2020, SongHJ2020, ChoiYS2020, HwangHS2020, KimHJ2021, YounJY2021, MinJW2022, KimJM2022, ShinHJ2023}.
    In general, when people understand speech or written text, they attempt to make sense of it using vocabulary and concepts that they are familiar with.
    While there are approaches to use rules or dictionaries to reflect this understanding~\cite{KwonHC1991}, it becomes difficult to build and maintain a dictionary for the vocabulary that appears in each text.
    Therefore, methods for tagging syllable units without a dictionary have been proposed~\cite{ShimKS2011, LeeCK2013, LeeCH2016, KimHM2016} and studies have been conducted to improve them ~\cite{KimSW2018, ChoiYS2018, KimHM2019, MinJW2019, SongHJ2019, SongHJ2020, YounJY2021, ShinHJ2023}.
    From a mechanical perspective, syllable-by-syllable morphological analysis can be performed either by tagging syllable-by-syllable and then applying a base-form restoration dictionary~\cite{ShimKS2011, LeeCH2016}, or by tagging syllable-by-syllable with the base form already restored~\cite{YounJY2021}.
    However, syllable-by-syllable morphological analysis has limitations, in that it is difficult to accurately identify morpheme boundaries and learn long-term contextual information as the length of the sequence increases.
    In this study, the former is referred to as dictionary-based morphological analysis and the latter as syllable-unit morphological analysis.
    Both methods are trained on a manually labeled corpus and cannot accurately analyze new syllable combinations or morphemes that do not appear in the training corpus.
    Recently, with the development of the Internet and the spread of open sources as well as open data, web texts, corpora, language resources, and knowledge shared by different people have accumulated significantly.
    The reduced cost of building and maintaining a dictionary provides a significant opportunity to overcome the limitations of dictionary-based methods.

    Against this background, this study considers how the dictionary-based morphological analysis method used by MeCab~\cite{MeCab}, an open software for Korean and Japanese morphological analysis in tokenizers, which is an essential preprocessing tool for deep learning, can be effectively improved, and a method is proposed.
    The dictionary-based morphological analysis method~\cite{Kudo2004, NaSH2014, NaSH2018} trained by the CRF (Conditional Random Fields) method~\cite{Lafferty2001} lists the candidate morphemes in the dictionary from a given sentence to form a lattice structure connected by a directed graph and determines the optimal morphological analysis path within it.
    The process of determining the optimal path in the lattice uses the Viterbi algorithm~\cite{Viterbi1967}, which determines the path that minimizes the cost of each morpheme node and the sum of the neighborhood costs of two consecutive morphemes.
    The main types of errors in these dictionary-based morphological analysis methods occur when new words that are not in the dictionary are used in a sentence, or when the optimal path calculation selects the incorrect result owing to bias.
    For example, it may be cost-effective to select one long morpheme than several short morphemes, but this may often lead to an incorrect analysis.
    The main motivation for this study was that the path that minimizes the costs for the nodes and links may not be the optimal path.

@@

    To identify cases in which a suboptimal solution is actually the best solution according to the best path calculation, we modified the best path calculation method to generate suboptimal analysis results and verified the extent to which they are correct.
    Although there are numerous different approaches to select the next-best path, we used the method of replacing a morpheme node on the optimal path with a lower-ranked node.
    As shown in Table~\ref{tab:maximum-performance}, we confirm the extent to which the analysis performance can be improved by replacing the optimal path with a lower-ranked node.
    We can consider the problem of finding the correct answer among the generated sub-optimals, similar to the problem of re-ranking search results in information retrieval~\cite{BaeYJ2021}.
    In ~\cite{ChoiYS2018}, the N-best analysis results generated by the seq2seq model were re-ranked based on a convolutional neural network to improve the performance.
    In this study, re-ranking was performed using two BERT models of different types and forms, as proposed in ~\cite{Nogueira2019}.
    Experimental results show that first-stage re-ranking improves the performance by over 20\% over existing written and spoken models, and second-stage re-ranking with a different type of input and a different type of pre-trained model further improves the performance by more than 30\% over existing written and spoken models.

    With this method, the performance of the dictionary-based morphological analysis method could be further improved; however, the overall analysis time increased when the morphological analysis system was configured, including the re-ranking model itself.
    However, it is feasible to use the results of multiple reranked morpheme analyses to update the connection costs between morphemes in a dictionary, similar to the backpropagation process in a typical neural network.
    It is also expected that the morphological analysis system with improved connection costs will be able to generate better re-ranking candidates, which will further improve performance by doing so iteratively.
    Further research is needed on this, and this study was limited to performance improvements with two-stage reranking.

    The main contributions of this study are as follows.
    \begin{enumerate}
        \item \textbf{Further improvement of dictionary-based morphological analysis method using suboptimal analysis results}: We explore the possibility of performance improvement by introducing a method to replace the optimal path with a suboptimal node and propose a method to effectively improve the dictionary-based morphological analysis method through deep learning.
        \item \textbf{Extending the performance improvement by introducing a two-stage re-ranking model}: To improve the performance of dictionary-based analysis by re-ranking the morphological analysis results, we propose extending the performance improvement using different BERT models to perform two rounds of re-ranking.
        \item \textbf{A method for updating connection costs in the dictionary and suggestions for future research}: We propose a new method for updating dictionary connection costs based on re-ranked morphological analysis results. We also outline directions for future research, suggesting potential improvements.
    \end{enumerate}
    These contributions provide important insights into the performance improvement of Korean morphological analysis and the direction of future research, and will serve as a useful reference for future researchers.

    The remainder of this paper is organized as follows.
    In Section ~\ref{sec:morphological-analysis-model}, we discuss configuring and training a dictionary-based morphological analysis system.
    In Section ~\ref{sec:reranking-model}, we discuss the generation of secondary results of morphological analysis, produce re-ranking data, and propose a method for training a two-stage re-ranking model.
    In Section ~\ref{sec:results}, we discuss the results of the performance improvement using the morphological analysis and re-ranking models.
    In Section ~\ref{sec:related-work}, we introduce previous research cases related to this study.
    Finally, in Section ~\ref{sec:conclusion}, we conclude the study and discuss its limitations and directions for future research.

>>>>

    Korean morphological analysis involves determining parts of speech by identifying morphemes, the smallest units of linguistic expression with independent meanings in a sentence.
    Unlike isolating languages like English, where sequential tagging suffices, Korean, being agglutinative, requires separating endings or postpositions and restoring inflections.
    The accuracy of morphological analysis significantly impacts Korean analysis performance since many tasks rely on separate morphemes as their basic input.
    Modern deep learning methods in natural language processing use tokenization, breaking text into smaller units and converting each into a vector for computational models~\cite{Mikolov2013}.
    For Korean, where subword units are crucial, attempting tokenization with separate morphemes in advance reflects the language's characteristics~\cite{SongHJ2021}.
    Incorporating morphological analysis results into this process enhances overall performance, capturing the semantic units of Korean.
    To accomplish this, we need a morphological analyzer that is not only highly accurate but also operates swiftly.

    Several approaches have been suggested for morphological analysis, a critical aspect of Korean language comprehension~\cite{KwonHC1991, LeeDG2009, ShimKS2011, LeeJS2011, ShinJC2012, LeeCK2013, NaSH2014, NaSH2015, HwangHS2016, KimHM2016, ChungES2016, LeeCH2016, Li2017, NaSH2018, KimSW2018, ChoiYS2018, MinJW2018, MinJW2019, KimHM2019, SongHJ2019, MinJW2020, SongHJ2020, ChoiYS2020, HwangHS2020, KimHJ2021, YounJY2021, MinJW2022, KimJM2022, ShinHJ2023}.
    Typically, when individuals grasp spoken or written language, they try to comprehend it through familiar vocabulary and concepts.
    While some approaches rely on rules or dictionaries to capture this understanding~\cite{KwonHC1991}, constructing and updating dictionaries for varied text vocabularies can be challenging.
    As a result, methods focusing on tagging syllable units without a dictionary have been proposed~\cite{ShimKS2011, LeeCK2013, LeeCH2016, KimHM2016} and studied for enhancement~\cite{KimSW2018, ChoiYS2018, KimHM2019, MinJW2019, SongHJ2019, SongHJ2020, YounJY2021, ShinHJ2023}.
    Mechanically, syllable-by-syllable morphological analysis can be achieved by either tagging syllables and then applying a base-form restoration dictionary~\cite{ShimKS2011, LeeCH2016} or by tagging syllables with the base form pre-restored~\cite{YounJY2021}.
    However, this approach has limitations, struggling with precise morpheme boundary identification and struggling to grasp long-term contextual information as the sequence lengthens.
    In this study, the former is termed dictionary-based morphological analysis, and the latter is syllable-unit morphological analysis.
    Both methods are trained on manually labeled corpora, facing challenges in accurately analyzing new syllable combinations or morphemes absent in the training data.
    The evolution of the Internet, open sources, and shared knowledge has led to substantial accumulations of web texts, corpora, language resources, offering an opportunity to overcome the constraints of dictionary-based methods due to reduced costs in dictionary construction and maintenance.

    Given this context, our study aims to enhance the effectiveness of the dictionary-based morphological analysis method employed by MeCab~\cite{MeCab}, an open-source tool for Korean and Japanese morphological analysis commonly used as a crucial preprocessing tool for deep learning.
    The method, trained through Conditional Random Fields (CRF)~\cite{Lafferty2001}, generates a lattice structure from a given sentence, connecting candidate morphemes in the dictionary through a directed graph.
    Subsequently, the optimal morphological analysis path is determined within this lattice structure~\cite{Kudo2004, NaSH2014, NaSH2018}.
    The Viterbi algorithm~\cite{Viterbi1967} is employed in this process, minimizing the cost associated with each morpheme node and the sum of neighborhood costs for consecutive morphemes to identify the optimal path.

    In these dictionary-based morphological analysis methods, the primary errors stem from encountering new words absent in the dictionary within a sentence or when biases lead to the selection of an incorrect result during optimal path calculation.
    For instance, opting for one long morpheme over several short ones might be cost-effective but often results in an inaccurate analysis.
    The main impetus behind our study is the recognition that the path minimizing costs for nodes and links may not always align with the optimal path.
    In response, we propose methods to address these challenges and improve the accuracy of the morphological analysis process.

@@

    To pinpoint instances where a suboptimal solution may, in fact, be the best choice according to the best path calculation, we modified the best path calculation method to yield suboptimal analysis results and assessed their accuracy.
    While various approaches exist for selecting the next-best path, we opted for the method of substituting a morpheme node on the optimal path with a lower-ranked node.
    Table~\ref{tab:maximum-performance} illustrates the degree to which analysis performance can be enhanced by replacing the optimal path with a lower-ranked node.
    This problem is analogous to the challenge of re-ranking search results in information retrieval~\cite{BaeYJ2021}, where the goal is to identify the correct answer among the generated suboptimal results.

    In ~\cite{ChoiYS2018}, the N-best analysis results produced by the seq2seq model were re-ranked based on a convolutional neural network to enhance performance.
    In our study, we employed re-ranking with two distinct BERT models, each of different types and forms, as proposed in ~\cite{Nogueira2019}.
    Experimental results reveal that first-stage re-ranking improves performance by over 20\% compared to existing written and spoken models.
    Furthermore, second-stage re-ranking, incorporating a different input type and a diverse pre-trained model, contributes to a performance improvement exceeding 30\% compared to existing written and spoken models.

    While our introduced method led to further enhancement in the performance of the dictionary-based morphological analysis, it resulted in an overall increase in analysis time when configuring the morphological analysis system, including the re-ranking model itself.
    However, a promising avenue for future exploration lies in utilizing the results of multiple re-ranked morpheme analyses to update the connection costs between morphemes in a dictionary, akin to the backpropagation process in a typical neural network.
    It is anticipated that an improved morphological analysis system with updated connection costs can generate superior re-ranking candidates, potentially enabling iterative performance improvements.
    While this study focused on two-stage re-ranking, further research is essential to fully explore this potential.

    The primary contributions of this study can be summarized as follows:
    \begin{enumerate}
        \item \textbf{Further improvement of dictionary-based morphological analysis method using suboptimal analysis results}: We investigate the potential for performance improvement by introducing a method to replace the optimal path with a suboptimal node. Additionally, we propose an effective approach to enhance the dictionary-based morphological analysis method through deep learning.
        \item \textbf{Extending the performance improvement by introducing a two-stage re-ranking model}: To further enhance the performance of dictionary-based analysis through re-ranking, we suggest extending the improvement using different BERT models and conducting two rounds of re-ranking.
        \item \textbf{A method for updating connection costs in the dictionary and suggestions for future research}: We present a novel method for updating dictionary connection costs based on re-ranked morphological analysis results. Furthermore, we outline directions for future research, suggesting potential enhancements.
    \end{enumerate}
    These contributions provide valuable insights into advancing the performance of Korean morphological analysis and offer guidance for future researchers.

    The subsequent sections of this paper are organized as follows:
    Section ~\ref{sec:morphological-analysis-model} discusses the configuration and training of a dictionary-based morphological analysis system.
    Section ~\ref{sec:reranking-model} covers the generation of secondary results of morphological analysis, the production of re-ranking data, and the proposal of a method for training a two-stage re-ranking model.
    Section ~\ref{sec:results} delves into the results of the performance improvement using morphological analysis and re-ranking models.
    Section ~\ref{sec:related-work} introduces previous research cases related to this study.
    Finally, in Section ~\ref{sec:conclusion}, we conclude the study, discuss its limitations, and suggest directions for future research.
